Data Exploration: IUCN and AquaMaps

### 2017-03-16 data_explore/data_explore.Rmd
done once and for all with this?

spp_class_all <- am_maps_class %>%
  full_join(iucn_maps_class, by = 'iucn_sid') %>%
  mutate(am_class = ifelse(am_class %in% c('null', 'not assigned'), NA, am_class),
         class    = if_else(is.na(am_class), iucn_class, am_class))

### Fix some missing class values
# spp_class_all1 %>% filter(is.na(class))
#       am_sid                   sciname am_class iucn_sid sciname_iucn iucn_class class
# 1  Fis-22442 Metapenaeopsis tarawensis     <NA>       NA         <NA>       <NA>  <NA> # Malacostraca
# 2 Fis-170350          Oenopota eriopis     <NA>       NA         <NA>       <NA>  <NA> # Gastropoda
# 3 Fis-170316          Yarica hyalosoma     <NA>       NA         <NA>       <NA>  <NA> # Actinopterygii
spp_class_all <- spp_class_all %>%
  mutate(class = case_when(
    is.na(.$class) & .$am_sid == 'Fis-22442' ~ 'Malacostraca',
    is.na(.$class) & .$am_sid == 'Fis-170350' ~ 'Gastropoda',
    is.na(.$class) & .$am_sid == 'Fis-170316' ~ 'Actinopterygii',
    TRUE ~ .$class))

@@

taxa_all <- taxa_am_all %>%
  full_join(taxa_am_solo,   by = 'class') %>%
  full_join(taxa_iucn_all,  by = 'class') %>%
  full_join(taxa_iucn_solo, by = 'class') %>%
  full_join(taxa_both,      by = 'class') %>%
  mutate(small_class = (is.na(n_iucn_class_all) & p_am_class_all < 0.5))

@@

class_to_kingdom <- read_csv(file.path(dir_M, 'git-annex/globalprep/_raw_data',
                                       'aquamaps/d2015/csv/speciesoccursum.csv')) %>%
  select(class, kingdom, phylum) %>%
  mutate(class = str_to_title(class)) %>%
  distinct()

taxa_other <- taxa_all %>%
  filter(small_class) %>%
  select(-small_class) %>%
  left_join(class_to_kingdom,
            by = 'class') %>%
  mutate(class = paste('other', str_to_title(kingdom), sep = ' ')) %>%
  select(-kingdom, -phylum) %>%
  group_by(class) %>%
  summarize(n_am_class_all = sum(n_am_class_all),
            p_am_class_all = sum(p_am_class_all),
            n_am_class_solo = sum(n_am_class_solo),
            p_am_class_solo = sum(p_am_class_solo))

@@

# print(bar_taxa)
ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
       width = 19.05, height = 7, units = 'cm', dpi = 600)

>>>>

spp_class_all <- am_maps_class %>%
  full_join(iucn_maps_class, by = 'iucn_sid') %>%
  mutate(am_class = ifelse(tolower(am_class) %in% c('null', 'not assigned'), NA, am_class),
         class    = ifelse(identical(iucn_class, am_class), am_class, NA),
         class    = ifelse(is.na(class & !is.na(iucn_class)), iucn_class, am_class))

### NAs due to null or not assigned Aquamaps species.
# spp_class_na <- spp_class_all %>%
#   filter(is.na(class)) %>%
#   select(am_sid, sciname)
# 
# spp_class_from_worrms <- data.table()
# for (i in seq_along(spp_class_na$sciname)) { ## i <- 1
#   am_sid  = spp_class_na$am_sid[i]
#   sciname = spp_class_na$sciname[i]
#   print(paste(i, sciname, sep = ' '))
#   x <- try(worrms::wm_records_taxamatch(name = spp_class_na$sciname[i]))
#   if(class(x) == 'try-error') {
#     x <- data.frame('am_sid'  = am_sid,
#                     'sciname' = sciname,
#                     'result'  = 'try-error')
#   } else {
#     x <- x %>%
#       as.data.frame() %>%
#       select(scientificname, valid_name, class, phylum, kingdom) %>%
#       mutate(sciname = sciname,
#              am_sid  = am_sid)
#   }
#   spp_class_from_worrms <- bind_rows(spp_class_from_worrms, x)
# }
# 
# ### still a few NAs:
# spp_class_from_worrms %>% filter(result == 'try-error') %>% select(am_sid, sciname)
# # Fis-170350	Oenopota eriopis			
# # Fis-170316	Yarica hyalosoma			
# # ITS-156736	Loxosoma murmanica
# spp_class_from_worrms1 <- spp_class_from_worrms %>%
#   mutate(class = case_when(
#     is.na(.$class) & .$am_sid == 'Fis-170350' ~ 'Gastropoda', ### from ITIS
#     is.na(.$class) & .$am_sid == 'Fis-170316' ~ 'Actinopterygii', ### from FishBase: cardinalfish
#     is.na(.$class) & .$am_sid == 'ITS-156736' ~ 'Entoprocta', ### from ITIS; Loxosomella murmanica
#     TRUE ~ .$class))
# write_csv(spp_class_from_worrms, file.path(dir_data, 'sciname_lookups/spp_class_from_worrms.csv'))

# spp_class_from_itis <- data.table()
# for (i in seq_along(spp_class_na$sciname)) { ## i <- 3
#   am_sid  = spp_class_na$am_sid[i]
#   sciname = spp_class_na$sciname[i]
#   print(paste(i, sciname, sep = ' '))
#   x <- taxize::classification(spp_class_na$sciname[i], 'itis')[[1]]
#   if(is.na(x)) {
#     y <- data.frame(am_sid  = am_sid,
#                   sciname = sciname,
#                   result = 'error')
#   } else {
#     class <- x$name[x$rank == 'class']
#     phylum <- x$name[x$rank == 'phylum']
#     kingdom <- x$name[x$rank == 'kingdom']
#     y <- data.frame(am_sid  = am_sid,
#                     sciname = sciname,
#                     class_itis = ifelse(length(class > 0), x$name[x$rank == 'class'], NA),
#                     phylum_itis = ifelse(length(phylum > 0), x$name[x$rank == 'phylum'], NA),
#                     kingdom_itis = ifelse(length(kingdom > 0), x$name[x$rank == 'kingdom'], NA))
#   }
#   spp_class_from_itis <- bind_rows(spp_class_from_itis, y)
# }
# write_csv(spp_class_from_itis, file.path(dir_data, 'sciname_lookups/spp_class_from_itis.csv'))

spp_class_fill <- read_csv(file.path(dir_data, 'sciname_lookups/spp_class_from_worrms.csv')) %>%
  select(-result) %>%
  full_join(read_csv(file.path(dir_data, 'sciname_lookups/spp_class_from_itis.csv')) %>%
                       select(-result),
            by = c('sciname', 'am_sid')) %>%
  mutate(class = ifelse(!is.na(class_worrms), class_worrms, class_itis)) %>%
  select(am_sid, sciname, class)
### NOTE: still some NAs; these are all class Animalia

spp_class_all <- spp_class_all %>%
  filter(!is.na(class)) %>%
  bind_rows(spp_class_fill)

@@

taxa_all <- taxa_am_all %>%
  full_join(taxa_am_solo,   by = 'class') %>%
  full_join(taxa_iucn_all,  by = 'class') %>%
  full_join(taxa_iucn_solo, by = 'class') %>%
  full_join(taxa_both,      by = 'class') %>%
  mutate(small_class = (is.na(n_iucn_class_all) & p_am_class_all < 0.5) | is.na(class))

@@

class_to_kingdom <- read_csv(file.path(dir_M, 'git-annex/globalprep/_raw_data',
                                       'aquamaps/d2015/csv/speciesoccursum.csv')) %>%
  select(class, kingdom, phylum) %>%
  bind_rows(read_csv(file.path(dir_data, 'sciname_lookups/spp_class_from_worrms.csv')) %>%
              select(class = class_worrms, kingdom = kingdom_worrms, phylum = phylum_worrms)) %>%
  bind_rows(read_csv(file.path(dir_data, 'sciname_lookups/spp_class_from_itis.csv')) %>%
              select(class = class_itis, kingdom = kingdom_itis, phylum = phylum_itis)) %>%
  mutate(class = str_to_title(class)) %>%
  select(-phylum) %>% ### avoid duplicates
  filter(!is.na(kingdom)) %>% ### NA class doesn't match with kingdom or phylum
  distinct() ### All remaining NAs are Animalia

taxa_other <- taxa_all %>%
  filter(small_class) %>%
  select(-small_class) %>%
  left_join(class_to_kingdom,
            by = 'class') %>%
  mutate(kingdom = ifelse(is.na(class), 'Animalia', kingdom),
         class = paste('other', str_to_title(kingdom), sep = ' ')) %>%
  select(-kingdom) %>%
  group_by(class) %>%
  summarize(n_am_class_all = sum(n_am_class_all),
            p_am_class_all = sum(p_am_class_all),
            n_am_class_solo = sum(n_am_class_solo),
            p_am_class_solo = sum(p_am_class_solo))

# taxa_other$n_am_class_all %>% sum(na.rm = TRUE) ### 968
# taxa_df$n_am_class_all %>% sum(na.rm = TRUE) ### 21921
### All AM species accounted for!

@@

print(bar_taxa)
ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), 
       width = 19.05, height = 7, units = 'cm', dpi = 600)
